---
title: Module 3 Notes
author: Justin Millar
date: '2018-09-13'
slug: module-3-notes
categories:
  - notes
tags: []
description: ''
---


#### [Download notes as a PDF](https://sta6093.netlify.com/pdf/2018-09-13-module-3-notes.pdf)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Upcoming Assignments


```{r, echo=FALSE, results='asis'}

t <- data.frame(
  Assignments = c("Module 4 Data Quiz", 
                  "Module 4 Conceptual Quiz"),
  "Open Time" = rep("Friday, Sept 14 (1:00 am EST)", 2),
  "Due Time"  = rep("Sunday, Sept 16 (11:55 pm EST)", 2), check.names = F
)

knitr::kable(t)
```

## Notes from Discussion Board/Office Hours

### Difference between `=` and `<-`

In R both `<-` and `=` can be used to assign values to variables/objects, however there are some slight differences in how R uses these operators behind-the-scenes. [This blog post](https://renkun.me/2014/01/28/difference-between-assignment-operators-in-r/) demonstrates the some these differences examples.

Typically I recommend always using `<-` for assigning variables, and I mostly reserve `=` for arguments within functions. A nice shortcut for <- is `Alt` + `-`.

```{r, eval=FALSE}
# My preferred uses of <- and =
q <- 0.12345
rounded_q <- round(q, digits = 0.2)
```



### Subsetting dataframes

#### Using the `subset()` function:

The `subset()` function can be used to select specific parts of a dataframe. It includes a `subset` argument for defining which rows will be kept, and a `select` argument for choosing which columns to include. Here's some examples using the `mtcarrs` data set:

```{r}
# Cars with more than 6 cylinders 
subset(mtcars, subset = cyl > 6)

# Cars with more than 6 cylinders and high miles per gallon
subset(mtcars, subset = cyl > 6 & hp > 150)

# Cars with more than 6 cylinders for high horsepower
subset(mtcars, subset = cyl > 6 | hp > 150)

# Just keep variables of interest
subset(mtcars, subset = cyl > 6 | hp > 100, select = c("cyl", "hp"))
```

#### Using operators:

We can also select part of a dataframe using brackets `[]`. Remember that this notation uses `[rows,columns]`to define a subset. It may be useful to also use the `$` operator, which most commonly used is to select a single column in a dataframe using the name of the column (as opposed to the index). This will return a vector, which may be useful for subsetting in an index (recall the Battleship example for the lecture). Here are the same examples as above just using the operators:

```{r}
# Cars with more than 6 cylinders 
mtcars[mtcars$cyl > 6,]

# Cars with more than 6 cylinders and high miles per gallon
mtcars[mtcars$cyl > 6 & mtcars$hp > 150,]

# Cars with more than 6 cylinders or high horsepower
mtcars[mtcars$cyl > 6 | mtcars$hp > 150,]

# Just keep variables of interest
mtcars[mtcars$cyl > 6 | mtcars$hp > 150, c("cyl", "hp")]

```

Both the these approaches will work the same, so use whichever you prefer. 


### Useful keyboard shortcuts

| Windows Shortcut: | Mac Shortcut: | Action: |
|---|---|---|
|`Ctrl` + `Enter`|`Cmd` + `Enter`| Run current line/selection |
|`Alt` + `Enter`|`Alt` + `Enter`| Run current line/selection (retain cursor position)|
|`Ctrl` + `Shift` + `Enter`|`Cmd` + `Shift` + `Enter`| Run Entire |
| `Alt` + `-`|`Alt` + `-`| Add `<-` |
| `Alt` + `Shift` + `K`|`Alt` + `Shift` + `K`| Show Shortcuts|

Full list of shortcuts [available here](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts?mobile_site=true)

### Extracting files from compressed (zipped) folders

In this course (and perhaps in your research) we will sometimes work with large files that must be compressed. Typically we will use .zip compression, which can be extracted by most newer PC and Mac without additional software. Other compression may require specific software to unzip (check the `Properties` of the folder, and try Google if you are unsure how to open). It is important that you either Extract the contents of the compressed folder, or manually move the files to a new location once to open the folder. This is because R will not allow you to specify a path to a file in a compressed folder. 


### Other notes
- [Cheat sheet for base R](https://www.rstudio.com/wp-content/uploads/2016/10/r-cheat-sheet-3.pdf)
- [List of other cheat sheets](https://rstudio.cloud/learn/cheat-sheets).
- Dataframes can be exported as CSV files using the `write.csv()` function.
- RStudio is available on [UFApps](https://info.apps.ufl.edu/).
- You can run RStudio on the cloud at https://rstudio.cloud/.
- [This video](https://www.youtube.com/watch?v=rVCBqg6xghI) demonstrates how to find and remove outliers using the `identify()` function.
-  Here's a [site with practice problems](http://rstatistics.net/).
